---
title: "Coding Assignment 3"
author: "Group 3"
date: "Due: 2023-12-09 23:59"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
#Put any packages you need here
knitr::opts_chunk$set(echo = TRUE)
```

A Florida health insurance company wants to predict annual claims for individual clients. The company pulls a random sample of 100 customers. The owner wishes to charge an actuarially fair premium to ensure a normal rate of return. The owner collects all of their current customer’s health care expenses from the last year and compares them with what is known about each customer’s plan. 

The data on the 100 customers in the sample is as follows:

-	Charges: Total medical expenses for a particular insurance plan (in dollars)
-	Age: Age of the primary beneficiary
-	BMI: Primary beneficiary’s body mass index (kg/m2)
-	Female: Primary beneficiary’s birth sex (0 = Male, 1 = Female)
-	Children: Number of children covered by health insurance plan (includes other dependents as well)
-	Smoker: Indicator if primary beneficiary is a smoker (0 = non-smoker, 1 = smoker)
-	Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all output, plots, etc. within this report.


```{r dataset, include=TRUE}
# # Install and load the readxl package
install.packages("readxl")
library(readxl)

# Read the data from an Excel file
insurance <- read.csv("../Data/Insurance_Data_Group3.csv")


```



## Question 1

Randomly select 30 observations from the sample and exclude from all modeling (i.e. n=47). Provide the summary statistics (min, max, std, mean, median) of the quantitative variables for the 70 observations.

```{r q1}
## Read the data from the CSV file
insurance <- read.csv("../Data/Insurance_Data_Group3.csv")

# Set seed for reproducibility
set.seed(123)

# Randomly select 30 observations to exclude
excluded_indices <- sample(1:nrow(insurance), 30)

# Select the remaining 70 observations
remaining_data <- insurance[-excluded_indices, ]

# Summary statistics
summary_stats <- summary(remaining_data[, c("Charges", "Age", "BMI")])
print(summary_stats)

```


## Question 2

Provide the correlation between all quantitative variables

```{r}
## Correlation matrix
correlation_matrix <- cor(remaining_data[, c("Charges", "Age", "BMI")])
print(correlation_matrix)

```


## Question 3

Run a regression that includes all independent variables in the data table. Does the model above violate any of the Gauss-Markov assumptions? If so, what are they and what is the solution for correcting?
```{r q3}
# Assuming 'Cities' is represented by dummy variables (e.g., City1, City2, ...)

# Run the regression excluding one city as the reference category
model <- lm(Charges ~ Age + BMI + Children + Female + Smoker + WinterSprings + WinterPark + Oviedo, data = insurance)

# Replace "..." with the rest of the dummy variables

# Check the summary of the model
summary(model)
```

## Question 4

Implement the solutions from question 3, such as data transformation, along with any other changes you wish. Use the sample data and run a new regression. How have the fit measures changed? How have the signs and significance of the coefficients changed?

```{r q4}
# Implement data transformations or other changes as needed
# For example, you might apply log transformations to the dependent or independent variables.

# Run a new regression
new_model <- lm(log(Charges) ~ Age + log(BMI) + Female + log(Children + 1) + Smoker + WinterSprings + WinterPark + Oviedo, data = remaining_data)

# Compare fit measures and coefficients with the previous model
summary(new_model)

```

## Question 5

Use the 30 withheld observations and calculate the performance measures for your best two models. Which is the better model? (remember that "better" depends on whether your outlook is short or long run)

```{r q5}
# Assuming your_data is your complete dataset
set.seed(123)  # For reproducibility

# Randomly select 30 observations to be withheld
withheld_indices <- sample(1:nrow(insurance), 30)

# Create a dataset with the withheld observations
withheld_data <- insurance[withheld_indices, ]

# The remaining data for training would be
training_data <- insurance[-withheld_indices, ]

# Assuming 'model' and 'new_model' are your regression models

# Predictions using the first model
predictions_model <- predict(model, newdata = withheld_data)

# Predictions using the second model
predictions_new_model <- predict(new_model, newdata = withheld_data)

# Calculate performance metrics for the first model
mse_model <- mean((predictions_model - withheld_data$Charges)^2)
rmse_model <- sqrt(mse_model)
r_squared_model <- 1 - (sum((withheld_data$Charges - predictions_model)^2) / sum((withheld_data$Charges - mean(withheld_data$Charges))^2))

# Repeat for the second model
mse_new_model <- mean((predictions_new_model - withheld_data$Charges)^2)
rmse_new_model <- sqrt(mse_new_model)
r_squared_new_model <- 1 - (sum((withheld_data$Charges - predictions_new_model)^2) / sum((withheld_data$Charges - mean(withheld_data$Charges))^2))

# Print or store the results
cat("Model 1 - MSE:", mse_model, "RMSE:", rmse_model, "R-squared:", r_squared_model, "\n")
cat("Model 2 - MSE:", mse_new_model, "RMSE:", rmse_new_model, "R-squared:", r_squared_new_model, "\n")


```


## Question 6

Provide interpretations of the coefficients, do the signs make sense? Perform marginal change analysis (thing 2) on the independent variables.

```{r}
# Assuming 'new_model' is the model you want to interpret

# Print the coefficients
coefficients <- coef(new_model)
print(coefficients)

# Interpretation of the coefficients
cat("Interpretation of Coefficients:\n")
cat("Intercept (Intercept):", coefficients[1], "\n")
cat("Age (Age):", coefficients[2], " - This represents the change in Charges for a one-unit change in Age.\n")
cat("BMI (log(BMI)):", coefficients[3], " - This represents the change in Charges for a one-unit change in log(BMI).\n")
cat("Female (Female):", coefficients[4], " - This represents the change in Charges for females compared to males (assuming binary gender).\n")
cat("Children (log(Children + 1)):", coefficients[5], " - This represents the change in Charges for a one-unit change in log(Children + 1).\n")
cat("Smoker (Smoker):", coefficients[6], " - This represents the change in Charges for smokers compared to non-smokers.\n")
cat("City (City2, City3, ...):", coefficients[7:length(coefficients)], " - Interpretation depends on the specific city dummy variables.\n")

# Perform marginal change analysis for a specific variable (e.g., Age)
marginal_change_age <- coefficients[2] * mean(remaining_data$Age)  # Replace 'mean(remaining_data$Age)' with your desired value
cat("\nMarginal Change in Charges for a one-unit change in Age:", marginal_change_age, "\n")

```

## Question 7

An eager insurance representative comes back with five potential clients. Using the better of the two models selected above, provide the prediction intervals for the five potential clients using the information provided by the insurance rep.

| Customer | Age | BMI | Female | Children | Smoker | City           |
| -------- | --- | --- | ------ | -------- | ------ | -------------- | 
| 1        | 60  | 22  | 1      | 0        | 0      | Oviedo         |
| 2        | 40  | 30  | 0      | 1        | 0      | Sanford        |
| 3        | 25  | 25  | 0      | 0        | 1      | Winter Park    |
| 4        | 33  | 35  | 1      | 2        | 0      | Winter Springs |
| 5        | 45  | 27  | 1      | 3        | 0      | Oviedo         |


```{r}
# Creating the potential_clients data frame
potential_clients <- data.frame(
  Customer = c(1, 2, 3, 4, 5),
  Age = c(60, 40, 25, 33, 45),
  BMI = c(22, 30, 25, 35, 27),
  Female = c(1, 0, 0, 1, 1),
  Children = c(0, 1, 0, 2, 3),
  Smoker = c(0, 0, 1, 0, 0),
  Oviedo = c(1, 0, 0, 0, 1),  # Replace with the correct dummy variables for each city
  Sanford = c(0, 1, 0, 0, 0),  # Replace with the correct dummy variables for each city
  WinterPark = c(0, 0, 1, 0, 0),  # Replace with the correct dummy variables for each city
  WinterSprings = c(0, 0, 0, 1, 0)  # Replace with the correct dummy variables for each city
)

# Display the potential_clients data frame
print(potential_clients)


# Assuming 'new_model' is the model you want to use
# Assuming 'potential_clients' is a data frame with information about the five potential clients

# Predictions for the potential clients with prediction intervals
predictions <- predict(new_model, newdata = potential_clients, interval = "prediction")

# Combine the original data and predictions
result <- cbind(potential_clients, predictions)
print(result)


```

## Question 8

The owner notices that some of the predictions are wider than others, explain why.
```{r q8}
# Assuming the 'better_model' is the model you want to analyze
residuals <- residuals(new_model)

# Plot residuals
plot(residuals)


```
The owner notices that some predictions are wider than others, the likely reason is varying prediction intervals. Prediction intervals widen when there is higher uncertainty associated with predictions. This can be influenced by factors like variability in the data or outliers.

Here's an explanation:

Wider Prediction Intervals: Predictions for some observations may have wider intervals when the model has higher uncertainty about the true value of the response variable.

Reasons for Wider Intervals:

Higher Variability in Data: If the data for certain observations are more variable, the model is less certain about the predictions.

Outliers: Outliers in the data can increase uncertainty, leading to wider intervals.

Non-linearity or Heteroscedasticity: If the relationship between predictors and response is non-linear or if the variability changes across the range of predictors, prediction intervals can be wider.


## Question 9 

Are there any prediction problems that occur with the five potential clients? If so, explain.
```{r q9}
# Assuming 'new_model' is the model you want to use
# Assuming 'potential_clients' is a data frame with information about the five potential clients

# Predictions for the potential clients with residuals
predictions <- predict(new_model, newdata = potential_clients, interval = "prediction")

# Extract residuals
residuals <- residuals(new_model, newdata = potential_clients)

# Combine the original data, predictions, and residuals
result <- cbind(potential_clients, predictions, residuals)
print(result)



```
The two regression models used are:

Model 1 (log-transformed Charges):

Model 1 MSE: 36481726
Model 1 RMSE: 6040.01
Model 1 R-squared: 0.7922652
Model 2 (original Charges):

Model 2 MSE: 416183189
Model 2 RMSE: 20400.57
Model 2 R-squared: -1.369837
Now, let's discuss potential prediction problems for the five potential clients:

Model Choice:

Model 1 (log-transformed Charges) seems to perform better, as it has a lower RMSE and a positive R-squared value, indicating a better fit.
Model 2, with a negative R-squared, suggests a poor fit and might not be suitable for predictions.
Prediction Interval Width:

Wider prediction intervals in Model 2 (original Charges) might be problematic for precise predictions.
Model 1 provides narrower intervals, indicating less uncertainty in predictions.
Outliers:

Both models have residuals, but Model 2 (original Charges) has a higher residual standard error, suggesting potential outliers affecting predictions.
Model Interpretability:

Model 1 coefficients are log-transformed, making them more interpretable in terms of percentage change.
Model 2 coefficients are in the original scale, which might be less interpretable.
R-squared:

Model 1 has a positive R-squared, suggesting it explains a reasonable proportion of the variance in the data.
Model 2 has a negative R-squared, indicating poor explanatory power.

Conclusion:
Considering the performance metrics and interpretability, Model 1 is preferred for predictions among the two models. However, it's crucial to note that prediction intervals are inherent in regression models, and predictions should be used cautiously, especially if the models are not perfect fits.